<<<<<<< HEAD
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}

# Unzip data to obtain a csv file

unzip("activity.zip")

# Read the data into activity data frame and show some summary statistics

data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}

# Make a histogram of the total number of steps taken each day

library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")

# Calculate and report the mean and median total number of steps taken per day

mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

```{r}
# Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
ylab("average number of steps taken")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

averages[which.max(averages$steps),]
```

## Imputing missing values

```{r}
# Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

# Calculate and report the total number of missing values in the dataset

missing <- is.na(data$steps)

table(missing)

# Fill in all of the missing values in the dataset with the mean value of its 5-minute interval

fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}

# Create a new dataset that is equal to the original dataset but with the missing data filled in

filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)

# Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day 

total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)

# These values differ from the estimates from the original dataset because because the missing values in the original dataset were set to 0 by default. But the new dataset have the filled-in missing values.
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)

# Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
xlab("5-minute interval") + ylab("Number of steps")
=======
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}

# Unzip data to obtain a csv file

unzip("activity.zip")

# Read the data into activity data frame and show some summary statistics

data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}

# Make a histogram of the total number of steps taken each day

library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")

# Calculate and report the mean and median total number of steps taken per day

mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

```{r}
# Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
ylab("average number of steps taken")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

averages[which.max(averages$steps),]
```

## Imputing missing values

```{r}
# Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

# Calculate and report the total number of missing values in the dataset

missing <- is.na(data$steps)

table(missing)

# Fill in all of the missing values in the dataset with the mean value of its 5-minute interval

fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}

# Create a new dataset that is equal to the original dataset but with the missing data filled in

filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)

# Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day 

total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)

# These values differ from the estimates from the original dataset because because the missing values in the original dataset were set to 0 by default. But the new dataset have the filled-in missing values.
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)

# Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
xlab("5-minute interval") + ylab("Number of steps")
>>>>>>> origin/master
```